Missing data: Current practice in football research and recommendations for improvement

### 2020-10-15 doc/eda.Rmd
from today's meeting

# Now we bring the third and final review 
gs_mismatches <- readr::read_csv("data/missing-data-mismatches.csv")

gs_mismatches %>% 
  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
               names_to = "reviewer",
               values_to = "decision") %>% 
  pivot_wider(names_from = variable,
              values_from = decision)

gs_football %>% 
  pivot_wider(names_from = reviewer,
              values_from = )

# supp materials - agreement between Reviewer 1 and Reviewer 2

#the agreement was `r agree_var(gs_football, was_data_shared)`%

>>>>

# Now we bring the third and final review 
gs_mismatches <- readr::read_csv("data/missing-data-mismatches.csv")
library(tidyverse)
gs_mismatches %>% 
  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
               names_to = "reviewer",
               values_to = "decision") %>% 
  pivot_wider(names_from = variable,
              values_from = decision)

# supp materials - agreement between Reviewer 1 and Reviewer 2
#the agreement was `r agree_var(gs_football, was_data_shared)`%

### 2020-10-15 paper/paper.Rmd
from today's meeting

```{r libraries, message = FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)
library(fitzRoy)
library(mice)
library(truncnorm)
library(janitor)
library(drake)
library(cowplot)
```

@@

It is imperative that missing data, or lack thereof, are reported [@Schafer2002]. 
While there has been significant interest in modelling missing
data [@refs], the exploration and reporting of missings have received
less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
of articles that report missing data in football research; (2) highlight
several practices that should be avoided; and (3) provide recommendations for
exploring and reporting missingness.

@@

To estimate the proportion of articles that report missing data in
football research, we conducted a systematic search [@PRISMA]. Articles on football topics (or involving footballers) published in 2019 were studied. Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key
terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process,
and further details of the search can also be found in Supplement 2.

```{r agree-var}
agree_var(gs_football, any_missing_data_statement)
agree_var(gs_football, any_statement_about_mcar_mnar_mar)
agree_var(gs_football, were_missing_data_explored)
agree_var(gs_football, was_imputation_used)
agree_var(gs_football, was_data_shared)

# identify table of mismatches  
gs_football %>% 
  group_by(reviewer) %>% 
  count(any_missing_data_statement)  %>% 
  pivot_wider(names_from = reviewer,
              values_from = n)
```

Our intention was to survey, at random, 10% of the total articles found (after 
title and abstract screening). We were primarily interested in
whether missing data was reported or acknowledged; and if not, whether
a dataset was shared. Where applicable, information relating to how missing
values were treated was also extracted (see Supplement 2). Two authors
independently extracted the data (X% agreement). Results are reported
as the proportion and 95% confidence interval.

@@

Of the 199 articles screened, 136 met the inclusion criteria, representing
\~10% of the articles found (see Supplement 2, Figure 1). The proportion of
articles that reported missing data was 5.9% (95% confidence interval;
X% to X%) or 8/136. Of the articles that did not mention missing data
(128/136), 7.1% (3.3% to 13.1%) or 9/128 shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration--especially in contrast
to many of the papers that included (potentially meaningless) tests of
normality (e.g., Kruskall-Wallis test).

>>>>

```{r libraries, message = FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)
library(fitzRoy)
library(mice)
library(truncnorm)
library(janitor)
library(drake)
library(scales)
```

@@

It is imperative that missing data, or lack thereof, are reported [@Schafer2002]. 
While there has been significant interest in modelling missing
data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received
less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
of articles that report missing data in football research; (2) highlight
several practices that should be avoided; and (3) provide recommendations for
exploring and reporting missingness.

@@

To estimate the proportion of articles that report missing data in
football research, we conducted a systematic search [@Moher2009]. Articles on football topics (or involving footballers) published in 2019 were studied. Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key
terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process,
and further details of the search can also be found in Supplement 2.

```{r agreement}

md_reported <- agree_var(gs_football, any_missing_data_statement) %>% 
  percent(scale = 1,
          accuracy = 0.1)

Our intention was to survey, at random, 10% of the total articles found (after 
title and abstract screening). We were primarily interested in
whether missing data was reported or acknowledged; and if not, whether
a dataset was shared. Where applicable, information relating to how missing
values were treated was also extracted (see Supplement 2). Two authors
independently extracted the data (`r md_reported` agreement). Results are reported
as the proportion and 95% confidence interval. (NOTE: need to get the final third reviewer part in)

@@

```{r}
# NOTE: update with the final third reviewer
# need to update the gs_football
library(binom)
# !!comment to explain NAs
successes <- sum(gs_football$any_missing_data_statement == 1, na.rm = TRUE)
trials <- sum(!is.na(gs_football$any_missing_data_statement))
missing_binom_conf <- binom.confint(x = successes, 
              n = trials, 
              conf.level = 0.95, 
              methods = "exact")

n_screened <- n_distinct(gs_football$key)
miss_reported_pct <- percent(missing_binom_conf$mean, accuracy = 0.1)
miss_reported_pct_lower <- percent(missing_binom_conf$lower, accuracy = 0.1)
miss_reported_pct_upper <- percent(missing_binom_conf$upper, accuracy = 0.1)

n_shared <- sum(gs_football$was_data_shared == 1)
```

```{r}
# now of the articles that don't mention missingness
gs_football_no_mention_sharing <- gs_football %>% 
  filter(any_missing_data_statement == 0)

n_no_mention_missings <- n_distinct(gs_football_no_mention_sharing$key)

no_mention_successes <- sum(gs_football_no_mention_sharing$was_data_shared == 1, 
                            na.rm = TRUE)
no_mention_trials <- sum(!is.na(gs_football_no_mention_sharing$was_data_shared))
n_mention_missing_binom_conf <- binom.confint(x = no_mention_successes, 
                                              n = no_mention_trials, 
                                              conf.level = 0.95, 
                                              methods = "exact")

percent_binom_conf <- function(binom_conf){
  mean_pct <- percent(n_mention_missing_binom_conf$mean, accuracy = 0.1)
lower_pct <- percent(n_mention_missing_binom_conf$lower, accuracy = 0.1)
upper_pct <- percent(n_mention_missing_binom_conf$upper, accuracy = 0.1)

return(c(mean_pct = mean_pct,
         lower_pct = lower_pct,
         upper_pct = upper_pct))
}

percent_binom_conf(n_mention_missing_binom_conf)

Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing
\~10% of the articles found (see Supplement 2, Figure 1). The proportion of
articles that reported missing data was `r miss_reported_pct` (95% confidence interval;
`r miss_reported_pct_lower` to `r miss_reported_pct_upper`) or `r successes`/`r n_screened`.

Of the articles that did not mention missing data `n_no_mention_missings` / `r n_screened`, 7.1% (3.3% to 13.1%) shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration--especially in contrast
to many of the papers that included (potentially meaningless) tests of
normality (e.g., Kruskall-Wallis test).
